Title of Paper

### 2017-02-16 paper.tex
update validation disc

In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
\begin{equation}
\frac{\text{d}[T]}{\text{d}t} = -\frac{\sum_{k}^{N_S} H^{\degree}_k \dot{\omega}_k}{\sum_{k}^{N_S} [C_k] c_{p_k}^{\degree}}
\end{equation}
although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.

@@

\section{Results and Discussion}
\subsection{Validation}
The reaction rates of progress, species and temperature rates in this study are validated by comparison to Cantera~\cite{Cantera}, however due to the intricacies of floating point arithmetic care must be taken when comparing results.

In obtaining the net rate of progress of reaction $i$ from the forward and reverse rates of progress, $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, a base-10 bound on the loss of precision can be found as~\cite{goldberg1991every}:
\begin{equation}
2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i}
\label{e:bounds}
\end{equation}
with $q_i \le p_i$, while $1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}}$ can be computed with accurately via use of higher-precision numbers (in this case, via use of Python's Decimal class).

Plugging the forward and reverse rate constants from Cantera into Eq.~\eqref{e:bounds} we can obtain a reasonable a-priori precision bounds $\text{P}_{\text{min},i} = 2^{-p_i}$, $\text{P}_{\text{max},i} = 2^{-q_i}$.
Finally, when determining the error in the net rates of progress between this code and Cantera, the range of percent error attributable to precision loss, $\text{P}_{\text{err},i}$ can be found as:
\begin{equation}
\text{P}_{i,\text{min}} = 100 * \frac{\text{P}_{\text{min},i}}{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert} \le
\text{P}_{i} \le 100 * \frac{\text{P}_{\text{max},i}}{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert} = \text{P}_{i,\text{max}}
\label{e:rel_bound}
\end{equation}
where the \text{CT} subscript indicates values from Cantera.
The error of a quantity $X_{i,j}$ for state $j$, and index $i$ was determined as:
\begin{equation}
 E_{X_{i}} = \left\lVert \left\lVert \frac{\left\lvert X_{i,j,\text{CT}} - X_{i,j}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{j,\infty} \right\rVert_{i, \infty}
\label{e:rel_err}
\end{equation}
Finally, to estimate the effect of loss of precision on the net rates of progress, we let $P_{\text{min}}^{*} = P_{i,\text{min}}$ for the $i,j$ value that results in $E_{R_{i}}$.

\begin{table}[ht]
\sisetup{retain-zero-exponent=true}
\centering
\begin{tabular}{@{}S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] @{}}
\toprule
\multicolumn{1}{c}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
\midrule
\multicolumn{1}{c}{\ce{H2}\slash~\ce{CO}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 4.66e6 & 2.37e1 & 2.12e5 \\
\multicolumn{1}{c}{GRI-Mech.~3.0}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 2.67e04 & 2.84e0 & 2.64e4 \\
\multicolumn{1}{c}{USC-Mech II}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
\bottomrule
\end{tabular}
\caption{Summary of rate of progress, species and temperature rate correctness.
Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
}
\label{T:error}
\end{table}

In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}

>>>>

In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
\begin{equation}
\frac{\text{d}[T]}{\text{d}t} = -\frac{\sum_{k}^{N_S} H^{\degree}_k \frac{\text{d}[C_k]}{\text{d}t}}{\sum_{k}^{N_S} [C_k] c_{p_k}^{\degree}}
\end{equation}
although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.

@@

\section{Results and Discussion}
\subsection{Validation}
The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues

In computing the net rate of progress of reaction $i$ from the forward and reverse ROP: $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, precision can easily be lost as the net rate of progress may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
A base-ten bound on loss of precision~\cite{goldberg1991every} can be found as:
\begin{equation}
\text{P}_{\text{min},i} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{P}_{\text{max},i}
\label{e:bounds}
\end{equation}
with $q_i \le p_i$.
To obtain an estimate of the a-priori precision bounds ($\text{P}_{\text{min},i}$ and $\text{P}_{\text{max},i}$), the forward and reverse ROP from Cantera are substituted into Eq.~\eqref{e:bounds}.
Additionally a higher-precision number format (in this case, the Python Decimal class) is used in this computation to maintain accuracy.
Finally, when determining the error in the net ROP between this code and Cantera, the bounds on the percent error attributable to precision loss, $\text{P}_{\text{err},i}$ can be estimated as:
\begin{equation}
\text{P}_{i,\text{min}} = 100 * \frac{\text{P}_{\text{min},i}}{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert} \le
\text{P}_{i} \le 100 * \frac{\text{P}_{\text{max},i}}{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert} = \text{P}_{i,\text{max}}
\label{e:rel_bound}
\end{equation}
where the \text{CT} subscript indicates values from Cantera.

For a more direct comparison, a relative error norm of a quantity $X_{i,j}$ over all states $j$, and indices $i$ was computed using the $L^{\infty}$ norm:
\begin{equation}
 E_{X_{i}} = \left\lVert \frac{\left\lvert X_{i,j,\text{CT}} - X_{i,j}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
\label{e:rel_err}
\end{equation}
and the lower precision bound for the net rate of progress error norm is set to $P_{\text{min}}^{*} = P_{i,\text{min}}$ for the $i, j$ value that results in the maximum error $E_{R_i}$.

\begin{table}[ht]
\sisetup{retain-zero-exponent=true}
\centering
\begin{tabular}{@{}S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] @{}}
\toprule
\multicolumn{1}{c}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
\midrule
\multicolumn{1}{c}{\ce{H2}\slash~\ce{CO}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 4.66e6 & 2.37e1 & 2.12e5 \\
\multicolumn{1}{c}{GRI-Mech.~3.0}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 2.67e04 & 2.84e0 & 2.64e4 \\
\multicolumn{1}{c}{USC-Mech II}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
\bottomrule
\end{tabular}
\caption{Summary of rate of progress, species and temperature rate correctness.
Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
}
\label{T:error}
\end{table}

In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
Very close agreement is observed for the forward and reverse ROP for all models, however the net rate of progress is \numrange{7}{8} orders of magnitude larger.
We see here that the minimum estimated loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining this jump.
The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera---as this loss of precision only occurs when the net ROP are many orders of magnitude smaller than the forward or reverse rates, implying the reaction is in near-equilibrium.
